---
title: "Figure 2: Differential Expression and Obesity"
output: html_notebook
---
```{r}
library(DESeq2)
library(dplyr)
library(biomaRt)
library(DEGreport)
library(ggplot2)
library(clusterProfiler)
library(ggpubr)
library(UpSetR)
library(cowplot)
library(tidyr)
library(ComplexHeatmap)
library(paletteer)
library(ggpmisc)
library(ReactomePA)
```

# Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set()
```

# Step 1: Import Data
Remember to set the appropriate working directory. The .RData file here was generated using "Data Wrangling" and "Preprocessing Scripts"
```{r}
load("./Longitudinal_complete_data.RData")
colfunc <- colorRampPalette(c("#832424", "white", "#3A3A98"))
```

## Set up Metadata
* Set "Day" as a factor - I want DESeq to consider nonlinear patterns in expression
* Add a bivariate, categorical classification for BMI, and make it a factor
* Set metadata rownames for downstream analyses
* Center BMI, seroconversion score, and age for DESeq
* Generate a table of different gene identifiers
```{r}
UGA4_longitudinal_metadata$Day <- factor(UGA4_longitudinal_metadata$Day)
UGA4_longitudinal_metadata$BMI_Bivariate <- ifelse(UGA4_longitudinal_metadata$BMI < 30, "Non_Obese",
                                                   "Obese")
UGA4_longitudinal_metadata$BMI_Bivariate <- factor(UGA4_longitudinal_metadata$BMI_Bivariate, levels = c("Non_Obese", "Obese"))

rownames(UGA4_longitudinal_metadata) <- paste0(UGA4_longitudinal_metadata$ID, "_", UGA4_longitudinal_metadata$Day)

UGA4_longitudinal_metadata$BMI_centered <- scale(UGA4_longitudinal_metadata$BMI, center = TRUE)
UGA4_longitudinal_metadata$seroconversion_centered <- scale(UGA4_longitudinal_metadata$vax_strain_seroconversion, center = TRUE)
UGA4_longitudinal_metadata$Age_centered <- scale(UGA4_longitudinal_metadata$Age, center = TRUE)

mart <- useEnsembl(biomart = "ensembl", "hsapiens_gene_ensembl")#, mirror = "useast")
geneids <- rownames(UGA4_longitudinal_counts_hemo_removed)
genenames <- as.data.frame(getBM(filters = "ensembl_gene_id",
                                 attributes = c("ensembl_gene_id", "hgnc_symbol", "entrezgene_id"),
                                 values = geneids,
                                 mart = mart))
```

## Set deseq import function
* Makes a function to make deseq objects from data, checks to make sure that the count colnames are equal to the metadata rownames, a requirement for deseq. Designed to be flexible when considering single day or multi-day data.
```{r}
deseq_import <- function(counts, metadata, design_setting = ~1) {
  ifelse(length(unique(metadata$Day)) > 1,
  rownames(metadata) <- paste0(metadata$ID, "_", metadata$Day),
  rownames(metadata) <- metadata$ID)

  if (all(rownames(metadata) == colnames(counts))) {
    object <- DESeqDataSetFromMatrix(
      countData = round(as.matrix(counts)),
      colData = metadata,
      design = design_setting)} else {
    "names not equal"
  }
  return(object)
}
```

## Divide data by day
* Divide data by days for analysis on individual days
```{r}
# Divide Day 0 Data
Day0_metadata <- UGA4_longitudinal_metadata %>% filter(Day == 0)
Day0_counts <- UGA4_longitudinal_counts_hemo_removed[colnames(UGA4_longitudinal_counts_hemo_removed) %in% rownames(Day0_metadata)]

# Divide Day 3 Data
Day3_metadata <- UGA4_longitudinal_metadata %>% filter(Day == 3)
Day3_counts <- UGA4_longitudinal_counts_hemo_removed[colnames(UGA4_longitudinal_counts_hemo_removed) %in% rownames(Day3_metadata)]
Day3_tpm <- UGA4_longitudinal_tpm_hemo_removed[colnames(UGA4_longitudinal_tpm_hemo_removed) %in% rownames(Day3_metadata)]

# Divide Day 7 Data
Day7_metadata <- UGA4_longitudinal_metadata %>% filter(Day == 7)
Day7_counts <- UGA4_longitudinal_counts_hemo_removed[colnames(UGA4_longitudinal_counts_hemo_removed) %in% rownames(Day7_metadata)]
Day7_tpm <- UGA4_longitudinal_tpm_hemo_removed[colnames(UGA4_longitudinal_tpm_hemo_removed) %in% rownames(Day7_metadata)]

# Divide Day 28 Data
Day28_metadata <- UGA4_longitudinal_metadata %>% filter(Day == 28)
Day28_counts <- UGA4_longitudinal_counts_hemo_removed[colnames(UGA4_longitudinal_counts_hemo_removed) %in% rownames(Day28_metadata)]
Day28_tpm <- UGA4_longitudinal_tpm_hemo_removed[colnames(UGA4_longitudinal_tpm_hemo_removed) %in% rownames(Day28_metadata)]
```

# Differential Expression by BMI
* Performs differential expression analysis with respect to BMI (BMI_centered)
* Keep only rows that have a count of at least 10 in the smallest group considered.
* Apply lfcshrink using the BMI variable.
* Pull out differentially expressed genes
```{r}
dds_d0 <- DESeqDataSetFromMatrix(
      countData = round(as.matrix(Day0_counts)),
      colData = Day0_metadata,
      design = ~ BMI_centered + Gender + Age_centered)

dds_d3 <- DESeqDataSetFromMatrix(
      countData = round(as.matrix(Day3_counts)),
      colData = Day3_metadata,
      design = ~ BMI_centered + Gender + Age_centered)

dds_d7 <- DESeqDataSetFromMatrix(
      countData = round(as.matrix(Day7_counts)),
      colData = Day7_metadata,
      design = ~ BMI_centered + Gender + Age_centered)

dds_d28 <- DESeqDataSetFromMatrix(
      countData = round(as.matrix(Day28_counts)),
      colData = Day28_metadata,
      design = ~ BMI_centered + Gender + Age_centered)

dds_d0 <- estimateSizeFactors(dds_d0)
keep_d0 <- rowSums(counts(dds_d0, normalized = TRUE) >= 10 ) >= 60
dds_d0 <- dds_d0[keep_d0, ]

dds_d3 <- estimateSizeFactors(dds_d3)
keep_d3 <- rowSums(counts(dds_d3, normalized = TRUE) >= 10 ) >= 60
dds_d3 <- dds_d3[keep_d3, ]

dds_d7 <- estimateSizeFactors(dds_d7)
keep_d7 <- rowSums(counts(dds_d7, normalized = TRUE) >= 10 ) >= 60
dds_d7 <- dds_d7[keep_d7, ]

dds_d28 <- estimateSizeFactors(dds_d28)
keep_d28 <- rowSums(counts(dds_d28, normalized = TRUE) >= 10 ) >= 60
dds_d28 <- dds_d28[keep_d28, ]

dds_d0 <- DESeq(dds_d0)

dds_d3 <- DESeq(dds_d3)

dds_d7 <- DESeq(dds_d7)

dds_d28 <- DESeq(dds_d28)

res_d0 <- lfcShrink(dds_d0, coef =  "BMI_centered")
res_d3<- lfcShrink(dds_d3, coef =  "BMI_centered")
res_d7 <- lfcShrink(dds_d7, coef = "BMI_centered")
res_d28 <- lfcShrink(dds_d28, coef =  "BMI_centered")

sig_d28 <- res_d28 %>%
              data.frame() %>%
              tibble::rownames_to_column(var = "gene") %>%
              as_tibble() %>%
              filter(padj < 0.05) %>%
              mutate(genename = genenames$hgnc_symbol[match(gene, genenames$ensembl_gene_id)])

sig_d7 <- res_d7 %>%
              data.frame() %>%
              tibble::rownames_to_column(var = "gene") %>%
              as_tibble() %>%
              filter(padj < 0.05) %>%
              mutate(genename = genenames$hgnc_symbol[match(gene, genenames$ensembl_gene_id)])

sig_d3 <- res_d3 %>%
              data.frame() %>%
              tibble::rownames_to_column(var = "gene") %>%
              as_tibble() %>%
              filter(padj < 0.05) %>%
              mutate(genename = genenames$hgnc_symbol[match(gene, genenames$ensembl_gene_id)])

sig_d0 <- res_d0 %>%
              data.frame() %>%
              tibble::rownames_to_column(var = "gene") %>%
              as_tibble() %>%
              filter(padj < 0.05) %>%
              mutate(genename = genenames$hgnc_symbol[match(gene, genenames$ensembl_gene_id)])
```

#Figure 2A
Generate an upset plot of DEGenes on each study day
```{r}
genes <- list("Day 0" = sig_d0$gene,
              "Day 3" = sig_d3$gene,
              "Day 7" = sig_d7$gene,
              "Day 28" = sig_d28$gene)
png("./OUTPUTS/Fig2A_Obesity_upset.png", height = 1250, width = 2000, res = 300, units = "px")
upset(fromList(genes[c(4, 3, 2, 1)]),
      keep.order = TRUE,
      sets = c("Day 28", "Day 7", "Day 3", "Day 0"),
      order.by = "freq",
      sets.x.label = "Number of\nDifferentially Expressed Genes",
      mainbar.y.label = "Number of Shared Genes",
      set_size.show = TRUE)
dev.off()

pdf("./OUTPUTS/Fig2A_Obesity_upset.pdf", height = 6, width = 8)
upset(fromList(genes[c(4, 3, 2, 1)]),
      keep.order = TRUE,
      sets = c("Day 28", "Day 7", "Day 3", "Day 0"),
      order.by = "freq",
      sets.x.label = "Number of\nDifferentially Expressed Genes",
      mainbar.y.label = "Number of Shared Genes",
      set_size.show = TRUE)
dev.off()

genes_with_intersection <- rbind(sig_d0 %>% mutate(Day = 0),
                                 sig_d3 %>% mutate(Day = 3),
                                 sig_d7 %>% mutate(Day = 7),
                                 sig_d28 %>% mutate(Day = 28)) %>%
  group_by(gene) %>%
  mutate(count = n()) %>%
  ungroup() %>%
  filter(count == 4)
```
# Export Differentially Expressed Genes
* Combine all tables of DEGenes into one and export for supplement
```{r}
all_degenes <- rbind(sig_d0 %>% mutate(Day = 0),
                                 sig_d3 %>% mutate(Day = 3),
                                 sig_d7 %>% mutate(Day = 7),
                                 sig_d28 %>% mutate(Day = 28))  %>%
   mutate(pvalue = signif(pvalue, 3),
          padj = signif(padj, 3),
          lfcSE = signif(lfcSE, 3),
          log2FoldChange = signif(log2FoldChange, 3),
          baseMean = signif(baseMean, 3))
write.csv(all_degenes, "./OUTPUTS/Supplement/Supplemental_Table_2.csv")
```

# Calculate enrichment per day: Gene lists
* Generates lists of genes for input into GSEA and Reactome enrichment functions. Removes NA/duplicated names from KEGG lists (used for Reactome)
```{r}
make_genelist <- function(results, KEGG = FALSE) {
  if(KEGG){
    list <- results$log2FoldChange
    names(list) <- rownames(results)
    names(list) <- genenames$entrezgene_id[match(names(list), genenames$ensembl_gene_id)]
  } else {
  list <- results$log2FoldChange
  names(list) <- rownames(results)
  }
  list <- sort(list, decreasing = TRUE)
    return(list)
}

genes_d0 <- make_genelist(res_d0)
genes_d0_kegg <- make_genelist(res_d0, KEGG = TRUE)
genes_d0_kegg <- genes_d0_kegg[!is.na(names(genes_d0_kegg))]
genes_d0_kegg <- genes_d0_kegg[!duplicated(names(genes_d0_kegg))]

genes_d3 <- make_genelist(res_d3)
genes_d3_kegg <- make_genelist(res_d3, KEGG = TRUE)
genes_d3_kegg <- genes_d3_kegg[!is.na(names(genes_d3_kegg))]
genes_d3_kegg <- genes_d3_kegg[!duplicated(names(genes_d3_kegg))]


genes_d7 <- make_genelist(res_d7)
genes_d7_kegg <- make_genelist(res_d7, KEGG = TRUE)
genes_d7_kegg <- genes_d7_kegg[!is.na(names(genes_d7_kegg))]
genes_d7_kegg <- genes_d7_kegg[!duplicated(names(genes_d7_kegg))]


genes_d28 <- make_genelist(res_d28)
genes_d28_kegg <- make_genelist(res_d28, KEGG = TRUE)
genes_d28_kegg <- genes_d28_kegg[!is.na(names(genes_d28_kegg))]
genes_d28_kegg <- genes_d28_kegg[!duplicated(names(genes_d28_kegg))]
```

# Calculate enrichment per day: Enrichment
* Calculates enriched terms for each study day, using a set seed for consistency of results.
* Generates a table of all enriched terms

```{r}
set.seed(139)
enrichment_d0 <- gseGO(genes_d0, 
                    "org.Hs.eg.db",
                    ont = "ALL",
                    keyType = "ENSEMBL",
                    #pvalueCutoff = 0.99,
                    pvalueCutoff = 0.05,
                    pAdjustMethod = "BH",
                    minGSSize = 10,
                    maxGSSize = 500,
                    eps = 0,
                    seed = TRUE)

enrichment_d3 <- gseGO(genes_d3, 
                    "org.Hs.eg.db",
                    ont = "ALL",
                    keyType = "ENSEMBL",
                    #pvalueCutoff = 0.99,
                    pvalueCutoff = 0.05,
                    pAdjustMethod = "BH",
                    minGSSize = 10,
                    maxGSSize = 500,
                    eps = 0,
                    seed = TRUE)

enrichment_d7 <- gseGO(genes_d7, 
                    "org.Hs.eg.db",
                    ont = "ALL",
                    keyType = "ENSEMBL",
                    #pvalueCutoff = 0.99,
                    pvalueCutoff = 0.05,
                    pAdjustMethod = "BH",
                    minGSSize = 10,
                    maxGSSize = 500,
                    eps = 0,
                    seed = TRUE)

enrichment_d28 <- gseGO(genes_d28, 
                    "org.Hs.eg.db",
                    ont = "ALL",
                    keyType = "ENSEMBL",
                    #pvalueCutoff = 0.99,
                    pvalueCutoff = 0.05,
                    pAdjustMethod = "BH",
                    minGSSize = 10,
                    maxGSSize = 500,
                    eps = 0,
                    seed = TRUE)

set.seed(139)
d0_reactome <- gsePathway(genes_d0_kegg,
                pvalueCutoff = 0.05,
                pAdjustMethod = "BH",
                verbose = FALSE,
                eps = 0,
                seed = TRUE)@result

d3_reactome <- gsePathway(genes_d3_kegg,
                pvalueCutoff = 0.05,
                pAdjustMethod = "BH",
                verbose = FALSE,
                eps = 0,
                seed = TRUE)@result

d7_reactome <- gsePathway(genes_d7_kegg,
                pvalueCutoff = 0.05,
                pAdjustMethod = "BH",
                verbose = FALSE,
                eps = 0,
                seed = TRUE)@result

d28_reactome <- gsePathway(genes_d28_kegg,
                pvalueCutoff = 0.05,
                pAdjustMethod = "BH",
                verbose = FALSE,
                eps = 0,
                seed = TRUE)@result

all_go <- rbind(enrichment_d0@result %>% mutate(Day = 0),
                enrichment_d3@result %>% mutate(Day = 3),
                enrichment_d7@result %>% mutate(Day = 7),
                enrichment_d28@result %>% mutate(Day = 28))

all_reactome <- rbind(d0_reactome %>% mutate(Day = 0),
                      d3_reactome %>% mutate(Day = 3),
                      d7_reactome %>% mutate(Day = 7),
                      d28_reactome %>% mutate(Day = 28))

all_terms <- rbind(all_go[, -1],
                   all_reactome)

```

# Export enrichment terms
* Generates a csv of all enriched terms for the supplement.
```{r}
all_terms_output <- all_terms %>%
   mutate(enrichmentScore = signif(enrichmentScore, 3),
          NES = signif(NES, 3),
          pvalue = signif(pvalue, 3),
          p.adjust = signif(p.adjust, 3),
          qvalue = signif(qvalue, 3))

write.csv(all_terms_output, "./OUTPUTS/Supplement/Supplemental_Table_3.csv", row.names = FALSE)
```
# Figure 2C: Tile Plot of Enriched Terms
* Plot curated terms in a tile plot.
```{r}
terms_to_plot <- c("R-HSA-6798695", "GO:0010818", "GO:0002385", "GO:0071396", "GO:0032607")

sig_sets <- all_terms %>% 
  group_by(ID) %>%
  mutate(n = matrixStats::count(p.adjust < 0.05)) %>%
  ungroup() %>%
  #filter(ONTOLOGY %in% c(NA, "BP")) %>% 
  #filter(n >= 2) %>%
  mutate(Signif = ifelse(p.adjust < 0.05, "*", "")) %>%
  arrange(n, Description)


terms_to_plot <- c("GO:0002237", "GO:0002385", "GO:0002251", "GO:0002227", "GO:0006959", "R-HSA-6798695", "GO:0019814", "GO:0071396", "GO:0042119", "GO:0150076", "GO:0030099", "GO:0061515", "R-HSA-168273", "R-HSA-168255")

png("./OUTPUTS/Fig2C_enrichment_tile.png", height = 1750, width = 2500, res = 300, units = "px")
ggplot(data = sig_sets %>% filter(ID %in% terms_to_plot), aes(x = factor(Day), y = Description, fill = NES)) +
  geom_tile() +
  geom_text(aes(label = Signif)) +
  labs(x = "Day", y = "") +
  scale_fill_gradient2() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 15, face = "bold"),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        element_text(family = "Helvetica"))
dev.off()

pdf("./OUTPUTS/Fig2C_enrichment_tile.pdf", height = 6, width = 8)
ggplot(data = sig_sets %>% filter(ID %in% terms_to_plot), aes(x = factor(Day), y = Description, fill = NES)) +
  geom_tile() +
  geom_text(aes(label = Signif)) +
  labs(x = "Day", y = "") +
  scale_fill_gradient2() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 15, face = "bold"),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        element_text(family = "Helvetica"))
dev.off()
```

# Figure 2B: Individual Gene Trends
* Plots individual genes to show trends of expression as a function of BMI
```{r}
genes_of_interest <- c("ENSG00000012223", "ENSG00000239839", "ENSG00000257017")
tpm_plots_genes_of_interest <- list()
for(i in 1:length(genes_of_interest)){
        plot_df <- data.frame(t(UGA4_longitudinal_tpm_hemo_removed[rownames(UGA4_longitudinal_tpm_hemo_removed) %in% genes_of_interest[[i]], ]))
        colnames(plot_df) <- genes_of_interest[[i]]
        plot_df$BMI <- UGA4_longitudinal_metadata$BMI[match(rownames(plot_df), rownames(UGA4_longitudinal_metadata))]
        plot_df$Day <- UGA4_longitudinal_metadata$Day[match(rownames(plot_df), rownames(UGA4_longitudinal_metadata))]

      tpm_plots_genes_of_interest[[i]] <- ggplot(data = plot_df, aes(x = !!plot_df[,2], y = !!plot_df[,1])) +
            geom_point(size = 1) +
            geom_smooth(method = "lm", color = "#07686f") +
            stat_fit_glance(method = 'lm',
                       method.args = list(formula = y ~ x),
                       geom = 'text',
                       aes(label = paste("P-value = ", signif(..p.value.., digits = 4), sep = "")),
                       label.x = 'middle', label.y = 'top', size = 5) +
            #scale_color_paletteer_d("nationalparkcolors::SmokyMountains", name = "BMI") +
            facet_grid(.~plot_df[,3]) +
            labs(title = genenames$hgnc_symbol[match(colnames(plot_df)[[1]], genenames$ensembl_gene_id)], x = "BMI", y = "TPM") +
            theme_bw() +
        scale_y_log10() +
            theme(axis.title.x = element_text(face = "bold", size = 20),
            axis.title.y = element_text(face = "bold", size = 20),
            axis.text = element_text(size = 15),
            title = element_text(size = 25),
            strip.background = element_blank(),
            strip.text = element_text(size = 15),
            aspect.ratio = 1,
            element_text(family = "Helvetica"))
}

tpm_plots_genes_of_interest

png("./OUTPUTS/Fig2B_ltf_tpm.png", height = 1250, width = 4000, res = 300, units = "px")
tpm_plots_genes_of_interest[[1]]
dev.off()

png("./OUTPUTS/Fig2B_DEFA3_tpm.png", height = 1250, width = 4000, res = 300, units = "px")
tpm_plots_genes_of_interest[[2]]
dev.off()

png("./OUTPUTS/Fig2B_HP_tpm.png", height = 1250, width = 4000, res = 300, units = "px")
tpm_plots_genes_of_interest[[3]]
dev.off()

pdf("./OUTPUTS/Fig2B_ltf_tpm.pdf", height = 7, width = 11)
tpm_plots_genes_of_interest[[1]]
dev.off()

pdf("./OUTPUTS/Fig2B_DEFA3_tpm.pdf", height = 7, width = 11)
tpm_plots_genes_of_interest[[2]]
dev.off()

pdf("./OUTPUTS/Fig2B_HP_tpm.pdf", height = 7, width = 11)
tpm_plots_genes_of_interest[[3]]
dev.off()
```