---
title: "Figure 5"
output: html_notebook
---

```{r}
library(tibble)
library(biomaRt)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggpmisc)
library(ppcor)
```

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/hockmanmr/OneDrive - National Institutes of Health/LAB_STUFF/PEOPLE/meg/manuscripts/CIVIC Longitudinal Transcriptomics")
```

# Step 1: Import Data
* Remember to set the appropriate working directory. The .RData file here was generated using "Data Wrangling" and "Preprocessing Scripts"
* This chunk also sets up some metadata covariates, and the rownames of the metadata.
* Additionally, generates a table of different gene identifiers that are more human-friendly
```{r}
#Load in data
load("./Scripts/Data/Publication/Longitudinal_complete_data.RData")

#Set up metadata covariates
UGA4_longitudinal_metadata$BMI_Bivariate <- ifelse(UGA4_longitudinal_metadata$BMI < 30, "Non-Obese",
                                                   "Obese")
UGA4_longitudinal_metadata$BMI_Bivariate <- factor(UGA4_longitudinal_metadata$BMI_Bivariate, levels = c("Non-Obese", "Obese"))
UGA4_longitudinal_metadata$Serology <- ifelse(UGA4_longitudinal_metadata$vax_strain_seroconversion <= 2,
                                                          "Non Responder", ifelse(UGA4_longitudinal_metadata$vax_strain_seroconversion < 4,
                                                          "Low Responder", "High Responder"))
UGA4_longitudinal_metadata$Serology <- factor(UGA4_longitudinal_metadata$Serology, levels = c("Non Responder", "Low Responder", "High Responder"))
UGA4_longitudinal_metadata$Day <- factor(UGA4_longitudinal_metadata$Day)
UGA4_longitudinal_metadata$n_prev_vax <- rowSums(UGA4_longitudinal_metadata[c("Vaccine.2.seasons.ago.(fall.2017-spring.2018)", "Vaccine.3.seasons.ago.(fall.2016-spring.2017)", "Vaccine.last.season.(fall.2018-spring.2019)")] == "Y")
UGA4_longitudinal_metadata$Age_categorical <- ifelse(UGA4_longitudinal_metadata$Age >= 65, "Elderly", "Young")

#Set metadata rownames
rownames(UGA4_longitudinal_metadata) <- paste0(UGA4_longitudinal_metadata$ID, "_", UGA4_longitudinal_metadata$Day)

#Make a table of human-readable gene names
mart <- useEnsembl(biomart = "ensembl", "hsapiens_gene_ensembl")#, mirror = "useast")
geneids <- rownames(UGA4_longitudinal_counts_hemo_removed)
genenames <- as.data.frame(getBM(filters = "ensembl_gene_id",
                                 attributes = c("ensembl_gene_id", "hgnc_symbol", "entrezgene_id"),
                                 values = geneids,
                                 mart = mart))
```

# Generate TPM table for cibersortx input
* This table will be input into the web cibersortx platform - the code only needs to be run once beforehand.
```{r}
UGA4_longitudinal_tpm_hemo_removed$symbol <- mapIds(org.Hs.eg.db,
                                                    keys = rownames(UGA4_longitudinal_tpm_hemo_removed),
                                                    column = "SYMBOL",
                                                    keytype = "ENSEMBL")

input_matrix <- UGA4_longitudinal_tpm_hemo_removed  %>%
  distinct(symbol, .keep_all = T) %>%
  rownames_to_column(var = "ensbl") %>%
  dplyr::select(-ensbl) %>%
  filter(!is.na(symbol)) %>%
  column_to_rownames(var = "symbol")

input_matrix <- na.omit(input_matrix)

input_matrix <- cbind(rownames(input_matrix), input_matrix)

colnames(input_matrix)[[1]] <- "GeneSymbols"

write.table(input_matrix, file = './Scripts/Data/cibersort_input_matrix_all_days.tsv', sep = "\t", 
            row.names = FALSE, col.names = TRUE, quote = FALSE)


write.table(mixture_file, file = "mixture_file.txt", sep = "\t", row.names = FALSE, col.names = TRUE,quote=FALSE)
```

# Generate Cibersort output tables
* This imports the cibersortx output tables that were downloaded from the internet runs, then makes tables for plot generation by adding metadata covariates of interest.
```{r}
cibersort_output_LM22 <- read.delim("./Scripts/Data/cibersort/CIBERSORTx_LM22.csv", sep = ",")
cibersort_output_MCCS <- read.delim("./Scripts/Data/cibersort/CIBERSORTx_MCCS.csv", sep = ",")
cibersort_output_immunoStates <- read.delim("./Scripts/Data/cibersort/CIBERSORTx_immunoStates.csv", sep = ",")

cell_analysis_LM22 <- cibersort_output_LM22 %>%
  mutate(Day = UGA4_longitudinal_metadata$Day[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         BMI = UGA4_longitudinal_metadata$BMI_Bivariate[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         BMI_Continuous = UGA4_longitudinal_metadata$BMI[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         Serology = UGA4_longitudinal_metadata$Serology[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         Serology_Continuous = log2(UGA4_longitudinal_metadata$vax_strain_seroconversion[match(rownames(UGA4_longitudinal_metadata), .$Mixture)]),
         n_prev_vax = UGA4_longitudinal_metadata$n_prev_vax[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         Age = UGA4_longitudinal_metadata$Age_categorical[match(rownames(UGA4_longitudinal_metadata), .$Mixture)])

cell_analysis_MCCS <- cibersort_output_MCCS %>%
  mutate(Day = UGA4_longitudinal_metadata$Day[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         BMI = UGA4_longitudinal_metadata$BMI_Bivariate[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         BMI_Continuous = UGA4_longitudinal_metadata$BMI[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         Serology = UGA4_longitudinal_metadata$Serology[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         Serology_Continuous = log2(UGA4_longitudinal_metadata$vax_strain_seroconversion[match(rownames(UGA4_longitudinal_metadata), .$Mixture)]),
         n_prev_vax = UGA4_longitudinal_metadata$n_prev_vax[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         Age = UGA4_longitudinal_metadata$Age_categorical[match(rownames(UGA4_longitudinal_metadata), .$Mixture)])

cell_analysis_immunoStates <- cibersort_output_immunoStates %>%
  mutate(Day = UGA4_longitudinal_metadata$Day[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         BMI = UGA4_longitudinal_metadata$BMI_Bivariate[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         BMI_Continuous = UGA4_longitudinal_metadata$BMI[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         Serology = UGA4_longitudinal_metadata$Serology[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         Serology_Continuous = log2(UGA4_longitudinal_metadata$vax_strain_seroconversion[match(rownames(UGA4_longitudinal_metadata), .$Mixture)]),
         n_prev_vax = UGA4_longitudinal_metadata$n_prev_vax[match(rownames(UGA4_longitudinal_metadata), .$Mixture)],
         Age = UGA4_longitudinal_metadata$Age_categorical[match(rownames(UGA4_longitudinal_metadata), .$Mixture)])
```

# Statistical Checks
* This is code to check the distributions of some of the data that will be compared, as well as compare groups of interest. It is meant to be edited based on groups of interest, and is not comprehensive as-written.
* The Shapiro test looks for normality of the data - if p > 0.05, we can assume the data are normal. If both groups are normally distributed, a t-test is appropriate. However, if one or both are non-normal, the wilcoxon test should be used. 
* As at least one of the comparison groups in the data are typically non-normally distributed, a wilcoxon test is used for comparisons.
```{r}
terms_to_plot <- c("CD8_positive_alpha_beta_T_cell", "T.cells.CD8", "DCs_IL_10", "Mast.cells.resting", "Plasma.cells", "B.cells.memory", "memory_B_cell")

variables <- c("BMI", "Serology")

Days <- c(0, 3, 7, 28)

unique(all_cibersort$Serology)

mean(all_cibersort %>% filter(BMI == 1) %>% pull(DCs_IL_10))
sd(all_cibersort %>% filter(BMI == 1) %>% pull(DCs_IL_10))
mean(all_cibersort %>% filter(BMI == 2) %>% pull(DCs_IL_10))
sd(all_cibersort %>% filter(BMI == 2) %>% pull(DCs_IL_10))


stats_data <- data.frame(matrix(NA, nrow = 56, ncol = 10))
mean_data <- data.frame(matrix(NA, nrow = ))

all_cibersort$BMI <- as.numeric(all_cibersort$BMI)
all_cibersort$Serology <- as.numeric(all_cibersort$Serology)

cell_analysis_MCCS$BMI_Stats <- ifelse(cell_analysis_MCCS$BMI == "Obese", 1, 0)
cell_analysis_MCCS$Serology

x <- as.data.frame(wilcox.test(all_cibersort %>%
              #filter(Serology == "Non Responder") %>% 
              #filter(BMI == "Non Obese") %>% 
              filter(Day == 3) %>%
              pull(Serology),
              all_cibersort %>% 
              #filter(Serology == "High Responder") %>% 
              #filter(BMI == 1) %>% 
              filter(Day == 3) %>%
              pull(DCs_IL_10),
            paired = FALSE))
 wilcox_effsize()

shapiro.test(cell_analysis_LM22 %>% 
         filter(Serology == "High Responder") %>% 
         filter(BMI == "Non Obese") %>% 
         filter(Day == 7) %>%
         pull(T.cells.CD8))

LM22_d3 <- cell_analysis_LM22 %>% filter(Day == 3) %>% filter(BMI == "Obese")

pcor.test(LM22_d3$Serology_Continuous, LM22_d3$T.cells.CD8 , LM22_d3$Age, method = "spearman")
cor.test(LM22_d3$Serology_Continuous, LM22_d3$T.cells.CD8, method = "spearman")

days <- c(0, 3, 5, 7)
lm22_cor <- data.frame(matrix(NA, nrow = (ncol(lm22_df) - 11)))
for(i in 1:4){
  lm22_df <- cell_analysis_LM22 %>% filter(Day == days[[i]])
  for(j in 2:(ncol(lm22_df) - 10)){
    lm22_cor[j, 1] <- colnames(lm22_df)[[j]]
    lm22_cor[j, 2] <- pcor.test(lm22_df[,j], lm22_df$BMI_Continuous, lm22_df$Age, method = "spearman")$p.value
  }
}

pcor.test(LM22_d3$BMI_Continuous, LM22_d3$B.cells.memory , LM22_d3$Age, method = "spearman")

cor.test(LM22_d7$BMI_Continuous, LM22_d7$B.cells.memory, method = "spearman")$p.value

spcor.test(immunoStates_d7$Serology_Continuous, immunoStates_d7$plasma_cell, immunoStates_d7$Age)
```

#Vaccine Response Comparison
* Plots all cell types stratified by vaccine response, for exploration.
```{r}
plotlist_LM22_vax <- list()
plotlist_LM22_vax_continuous <- list()

plotlist_MCCS_vax <- list()
plotlist_MCCS_vax_continuous <- list()

plotlist_immunoStates_vax <- list()
plotlist_immunoStates_vax_continuous <- list()

pops_LM22 <- colnames(cell_analysis_LM22)[c(2:23)]
pops_MCCS <- colnames(cell_analysis_MCCS)[c(2:13)]
pops_immunoStates <- colnames(cell_analysis_immunoStates)[c(2:21)]

for(i in 1:length(pops_LM22)){
 plotlist_LM22_vax[[i]] <- ggplot(cell_analysis_LM22, aes(x = Serology, y = !!cell_analysis_LM22[, pops_LM22[[i]]], color = Serology)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#601A4A", "#EE442F", "#63ACBE")) +
  labs(title = pops_LM22[[i]]) +
   geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
   facet_grid(.~Day) +
   theme_bw()
}

for(i in 1:length(pops_LM22)){
 plotlist_LM22_vax_continuous[[i]] <- ggplot(cell_analysis_LM22, aes(x = Serology_Continuous, y = !!cell_analysis_LM22[, pops_LM22[[i]]])) +
  geom_smooth(method = "lm", color = "#9D5557") +
  geom_point(alpha = 0.5) +
  stat_fit_glance(method = 'lm',
                       method.args = list(formula = y ~ x),
                       geom = 'text',
                       aes(label = paste("P-value = ", signif(..p.value.., digits = 4), sep = "")),
                       label.x = 'middle', label.y = 'top', size = 5) +
  labs(title = pops_LM22[[i]]) +
  facet_grid(.~Day) +
  theme_bw()
}

for(i in 1:length(pops_MCCS)){
 plotlist_MCCS_vax[[i]] <- ggplot(cell_analysis_MCCS, aes(x = Serology, y = !!cell_analysis_MCCS[, pops_MCCS[[i]]], color = Serology)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#601A4A", "#EE442F", "#63ACBE")) +
  labs(title = pops_MCCS[[i]]) +
   geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
   facet_grid(.~Day) +
   theme_bw()
}

for(i in 1:length(pops_MCCS)){
 plotlist_MCCS_vax_continuous[[i]] <- ggplot(cell_analysis_MCCS, aes(x = Serology_Continuous, y = !!cell_analysis_MCCS[, pops_MCCS[[i]]])) +
  geom_smooth(method = "lm", color = "#9D5557") +
  geom_point(alpha = 0.5) +
  stat_fit_glance(method = 'lm',
                       method.args = list(formula = y ~ x),
                       geom = 'text',
                       aes(label = paste("P-value = ", signif(..p.value.., digits = 4), sep = "")),
                       label.x = 'middle', label.y = 'top', size = 5) +
  labs(title = pops_MCCS[[i]]) +
  facet_grid(.~Day) +
  theme_bw()
}

for(i in 1:length(pops_immunoStates)){
 plotlist_immunoStates_vax[[i]] <- ggplot(cell_analysis_immunoStates, aes(x = Serology, y = !!cell_analysis_immunoStates[, pops_immunoStates[[i]]], color = Serology)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#601A4A", "#EE442F", "#63ACBE")) +
  labs(title = pops_immunoStates[[i]]) +
   geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
   facet_grid(.~Day) +
   theme_bw()
}

for(i in 1:length(pops_immunoStates)){
 plotlist_immunoStates_vax_continuous[[i]] <- ggplot(cell_analysis_immunoStates, aes(x = Serology_Continuous, y = !!cell_analysis_immunoStates[, pops_immunoStates[[i]]])) +
  geom_smooth(method = "lm", color = "#9D5557") +
  geom_point(alpha = 0.5) +
  stat_fit_glance(method = 'lm',
                       method.args = list(formula = y ~ x),
                       geom = 'text',
                       aes(label = paste("P-value = ", signif(..p.value.., digits = 4), sep = "")),
                       label.x = 'middle', label.y = 'top', size = 5) +
  labs(title = pops_immunoStates[[i]]) +
  facet_grid(.~Day) +
  theme_bw()
}

plotlist_immunoStates_vax
plotlist_MCCS_vax
plotlist_LM22_vax
plotlist_LM22_vax_continuous
plotlist_immunoStates_vax_continuous
plotlist_MCCS_vax_continuous
```

#BMI Comparison
* Plots all cell types stratified by BMI, for exploration.
```{r}
plotlist_LM22_bmi <- list()
plotlist_LM22_bmi_continuous <- list()

plotlist_MCCS_bmi <- list()
plotlist_MCCS_bmi_continuous <- list()

plotlist_immunoStates_bmi <- list()
plotlist_immunoStates_bmi_continuous <- list()

pops_LM22 <- colnames(cell_analysis_LM22)[c(2:23)]
pops_MCCS <- colnames(cell_analysis_MCCS)[c(2:13)]
pops_immunoStates <- colnames(cell_analysis_immunoStates)[c(2:21)]

for(i in 1:length(pops_LM22)){
 plotlist_LM22_bmi[[i]] <- ggplot(cell_analysis_LM22, aes(x = Day, y = !!cell_analysis_LM22[, pops_LM22[[i]]], color = BMI)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#289F90", "#EAC36F")) +
  labs(title = pops_LM22[[i]]) +
   geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
   #facet_grid(n_prev_vax~.) +
   theme_bw()
}

plotlist_LM22_bmi[[1]]

for(i in 1:length(pops_LM22)){
 plotlist_LM22_bmi_continuous[[i]] <- ggplot(cell_analysis_LM22, aes(x = BMI_Continuous, y = !!cell_analysis_LM22[, pops_LM22[[i]]])) +
  geom_smooth(method = "lm", color = "#9D5557") +
  geom_point(alpha = 0.5) +
  stat_fit_glance(method = 'lm',
                       method.args = list(formula = y ~ x),
                       geom = 'text',
                       aes(label = paste("P-value = ", signif(..p.value.., digits = 4), sep = "")),
                       label.x = 'middle', label.y = 'top', size = 5) +
  labs(title = pops_LM22[[i]]) +
  facet_grid(.~Day) +
  theme_bw()
}

for(i in 1:length(pops_MCCS)){
 plotlist_MCCS_bmi[[i]] <- ggplot(cell_analysis_MCCS, aes(x = Day, y = !!cell_analysis_MCCS[, pops_MCCS[[i]]], color = BMI)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#289F90", "#EAC36F")) +
  labs(title = pops_MCCS[[i]]) +
   geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
   #facet_grid(n_prev_vax~.) +
   theme_bw()
}

for(i in 1:length(pops_MCCS)){
 plotlist_MCCS_bmi_continuous[[i]] <- ggplot(cell_analysis_MCCS, aes(x = BMI_Continuous, y = !!cell_analysis_MCCS[, pops_MCCS[[i]]])) +
  geom_smooth(method = "lm", color = "#9D5557") +
  geom_point(alpha = 0.5) +
  stat_fit_glance(method = 'lm',
                       method.args = list(formula = y ~ x),
                       geom = 'text',
                       aes(label = paste("P-value = ", signif(..p.value.., digits = 4), sep = "")),
                       label.x = 'middle', label.y = 'top', size = 5) +
  labs(title = pops_MCCS[[i]]) +
  facet_grid(.~Day) +
  theme_bw()
}

for(i in 1:length(pops_immunoStates)){
 plotlist_immunoStates_bmi[[i]] <- ggplot(cell_analysis_immunoStates, aes(x = Day, y = !!cell_analysis_immunoStates[, pops_immunoStates[[i]]], color = BMI)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#289F90", "#EAC36F")) +
  labs(title = pops_immunoStates[[i]]) +
   geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
   #facet_grid(n_prev_vax~Day) +
   theme_bw()
}

for(i in 1:length(pops_immunoStates)){
 plotlist_immunoStates_bmi_continuous[[i]] <- ggplot(cell_analysis_immunoStates, aes(x = BMI_Continuous, y = !!cell_analysis_immunoStates[, pops_immunoStates[[i]]])) +
  geom_smooth(method = "lm", color = "#9D5557") +
  geom_point(alpha = 0.5) +
  stat_fit_glance(method = 'lm',
                       method.args = list(formula = y ~ x),
                       geom = 'text',
                       aes(label = paste("P-value = ", signif(..p.value.., digits = 4), sep = "")),
                       label.x = 'middle', label.y = 'top', size = 5) +
  labs(title = pops_immunoStates[[i]]) +
  facet_grid(.~Day) +
  theme_bw()
}

plotlist_immunoStates_bmi
plotlist_MCCS_bmi
plotlist_LM22_bmi
#plotlist_LM22_bmi_continuous
#plotlist_immunoStates_bmi_continuous
#plotlist_MCCS_bmi_continuous
```

#Bivariate Comparison
* Plots all cell types stratified by vaccine response and BMI, for exploration.
```{r}
plotlist_LM22_bivariate <- list()

plotlist_MCCS_bivariate <- list()

plotlist_immunoStates_bivariate <- list()

pops_LM22 <- colnames(cell_analysis_LM22)[c(2:23)]
pops_MCCS <- colnames(cell_analysis_MCCS)[c(2:13)]
pops_immunoStates <- colnames(cell_analysis_immunoStates)[c(2:21)]

for(i in 1:length(pops_LM22)){
 plotlist_LM22_bivariate[[i]] <- ggplot(cell_analysis_LM22, aes(x = Day, y = !!cell_analysis_LM22[, pops_LM22[[i]]], color = Serology)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#601A4A", "#EE442F", "#63ACBE")) +
  labs(title = pops_LM22[[i]]) +
   geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
   facet_grid(.~BMI) +
   theme_bw()
}

for(i in 1:length(pops_MCCS)){
 plotlist_MCCS_bivariate[[i]] <- ggplot(cell_analysis_MCCS, aes(x = Day, y = !!cell_analysis_MCCS[, pops_MCCS[[i]]], color = Serology)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#601A4A", "#EE442F", "#63ACBE")) +
  labs(title = pops_MCCS[[i]]) +
   geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
   facet_grid(.~BMI) +
   theme_bw()
}

for(i in 1:length(pops_immunoStates)){
 plotlist_immunoStates_bivariate[[i]] <- ggplot(cell_analysis_immunoStates, aes(x = Day, y = !!cell_analysis_immunoStates[, pops_immunoStates[[i]]], color = Serology)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#601A4A", "#EE442F", "#63ACBE")) +
  labs(title = pops_immunoStates[[i]]) +
   geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
   facet_grid(.~BMI) +
   theme_bw()
}


plotlist_immunoStates_bivariate
plotlist_MCCS_bivariate
plotlist_LM22_bivariate
```

```{r}
terms_to_plot <- c("Plasma.cells", "T.cells.CD8", "DCs_IL_10")

serology_comparisons <- list(c("High Responder", "Non Responder"), c("High Responder", "Low Responder"), c("Low Responder", "Non Responder"))

all_cibersort <- merge(cell_analysis_immunoStates, cell_analysis_LM22, by = c("BMI", "BMI_Continuous", "Serology", "Serology_Continuous", "Day", "n_prev_vax", "Age"))
all_cibersort <- merge(all_cibersort, cell_analysis_MCCS, by = c("BMI", "BMI_Continuous", "Serology", "Serology_Continuous", "Day", "n_prev_vax", "Age"))

all_cibersort$Day <- factor(all_cibersort$Day)

plotlist_figure_a <- list()
plotlist_figure_b <- list()
plotlist_figure_c <- list()

for(i in 1:length(terms_to_plot)){
 plotlist_figure_a[[i]] <- ggplot(all_cibersort, aes(x = Day, y = !!all_cibersort[, terms_to_plot[[i]]], color = Serology)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#601A4A", "#EE442F", "#63ACBE")) +
  labs(title = terms_to_plot[[i]], y = "Proportion") +
  geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
  theme_bw() +
  theme(strip.background = element_blank(),
        axis.text.x = element_text(face = "bold",size = 12),
aspect.ratio = 1,
axis.title = element_text(size = 14),
strip.text = element_text(size = 14),
legend.title = element_text(face = "bold", size = 12),
title = element_blank())
 
plotlist_figure_b[[i]] <- ggplot(all_cibersort, aes(x = Day, y = !!all_cibersort[, terms_to_plot[[i]]], color = BMI)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#289F90", "#EAC36F")) +
  labs(title = terms_to_plot[[i]], y = "Proportion") +
  geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
  theme_bw() +
  theme(strip.background = element_blank(),
        axis.text.x = element_text(face = "bold", size = 12),
aspect.ratio = 1,
axis.title = element_text(size = 14),
strip.text = element_text(size = 14),
legend.title = element_text(face = "bold", size = 12),
title = element_blank())

 plotlist_figure_c[[i]] <- ggplot(all_cibersort, aes(x = Day, y = !!all_cibersort[, terms_to_plot[[i]]], color = Serology)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#601A4A", "#EE442F", "#63ACBE")) +
  labs(title = terms_to_plot[[i]], y = "Proportion") +
  geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
  facet_grid(~BMI) +
  theme_bw() +
  theme(strip.background = element_blank(),
        axis.text.x = element_text(face = "bold",
size = 12),
aspect.ratio = 1,
axis.title = element_text(size = 14),
strip.text = element_text(size = 14),
legend.title = element_text(face = "bold", size = 12),
title = element_blank())
}

plotlist_figure_a
plotlist_figure_b
plotlist_figure_c
for(i in 1:length(plotlist_figure_a)){
  pdf(paste0("./Figures/Fig5/",terms_to_plot[[i]], "_response.pdf"))
  print(plotlist_figure_a[[i]])
  dev.off()
}

for(i in 1:length(plotlist_figure_b)){
  pdf(paste0("./Figures/Fig5/",terms_to_plot[[i]], "_bmi.pdf"))
  print(plotlist_figure_b[[i]])
  dev.off()
}

for(i in 1:length(plotlist_figure_c)){
  pdf(paste0("./Figures/Fig5/",terms_to_plot[[i]], "_bmi_response.pdf"))
  print(plotlist_figure_c[[i]])
  dev.off()
}

for(i in 1:length(plotlist_figure_b)){
  png(filename = paste0("./Figures/Fig5/",terms_to_plot[[i]], "_bmi.png"), height = 1600, width = 1700, res = 300, units = "px")
  print(plotlist_figure_b[[i]])
  dev.off()
}

for(i in 1:length(plotlist_figure_c)){
  png(filename = paste0("./Figures/Fig5/",terms_to_plot[[i]], "_bmi_response.png"), height = 1600, width = 3000, res = 300, units = "px")
  print(plotlist_figure_c[[i]])
  dev.off()
}
```

```{r}
terms_to_plot <- c("T.cells.CD8", "DCs_IL_10", "Mast.cells.resting", "Plasma.cells", "B.cells.memory", "memory_B_cell")

all_cibersort <- merge(cell_analysis_immunoStates, cell_analysis_LM22, by = c("BMI", "BMI_Continuous", "Serology", "Serology_Continuous", "Day", "n_prev_vax", "Age"))
all_cibersort <- merge(all_cibersort, cell_analysis_MCCS, by = c("BMI", "BMI_Continuous", "Serology", "Serology_Continuous", "Day", "n_prev_vax", "Age"))

all_cibersort$Day <- factor(all_cibersort$Day)

plotlist_supplement_a <- list()
plotlist_supplement_b <- list()
plotlist_supplement_c <- list()

for(i in 1:length(terms_to_plot)){
 plotlist_supplement_a[[i]] <- ggplot(all_cibersort, aes(x = Day, y = !!all_cibersort[, terms_to_plot[[i]]], color = Serology)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#601A4A", "#EE442F", "#63ACBE")) +
  labs(title = terms_to_plot[[i]], y = "Proportion") +
  geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
  facet_grid(.~Age) +
  theme_bw() +
  theme(strip.background = element_blank(),
        axis.text.x = element_text(face = "bold",size = 12),
aspect.ratio = 1,
axis.title = element_text(size = 14),
strip.text = element_text(size = 14),
legend.title = element_text(face = "bold", size = 12))
#title = element_blank())
 
plotlist_supplement_b[[i]] <- ggplot(all_cibersort, aes(x = Day, y = !!all_cibersort[, terms_to_plot[[i]]], color = BMI)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#289F90", "#EAC36F")) +
  labs(title = terms_to_plot[[i]], y = "Proportion") +
  facet_grid(.~Age) +
  geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
  theme_bw() +
  theme(strip.background = element_blank(),
        axis.text.x = element_text(face = "bold", size = 12),
aspect.ratio = 1,
axis.title = element_text(size = 14),
strip.text = element_text(size = 14),
legend.title = element_text(face = "bold", size = 12))
#title = element_blank())

 plotlist_supplement_c[[i]] <- ggplot(all_cibersort, aes(x = Day, y = !!all_cibersort[, terms_to_plot[[i]]], color = Serology)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), alpha = 0.5) +
  scale_color_manual(values = c("#601A4A", "#EE442F", "#63ACBE")) +
  labs(title = terms_to_plot[[i]], y = "Proportion") +
   geom_pwc(method = "wilcox_test", label = "p.signif", hide.ns = TRUE, vjust = 0.5) +
  facet_grid(Age~BMI) +
  theme_bw() +
  theme(strip.background = element_blank(),
        axis.text.x = element_text(face = "bold",
size = 12),
aspect.ratio = 1,
axis.title = element_text(size = 14),
strip.text = element_text(size = 14),
legend.title = element_text(face = "bold", size = 12))
#title = element_blank())
}

plotlist_supplement_a
plotlist_supplement_b
plotlist_supplement_c

pdf("./Figures/Supplement/Fig5Supplement.pdf", onefile = TRUE)
#Plasma Cells
plotlist_supplement_a[[4]]
plotlist_supplement_c[[4]]
#IL-10 DCs
plotlist_supplement_b[[2]]
plotlist_supplement_c[[2]]
#CD8 T Cells
plotlist_supplement_c[[1]]
dev.off()
```

###DEPRECATED#######

```{r}
terms_to_plot <- c("CD8_positive_alpha_beta_T_cell", "T.cells.CD8", "DCs_IL_10", "Mast.cells.resting", "Plasma.cells", "B.cells.memory", "memory_B_cell")
vars_3 <- c("Serology", "BMI", "Day")
vars_2 <- c("Serology", "Day")

all_cibersort <- merge(cell_analysis_immunoStates[, !colnames(cell_analysis_immunoStates) %in% c("P.value", "Correlation", "RMSE")], 
                       cell_analysis_LM22[, !colnames(cell_analysis_LM22) %in% c("P.value", "Correlation", "RMSE")], 
                       by = c("BMI", "BMI_Continuous", "Serology", "Serology_Continuous", "Day", "n_prev_vax", "Age"))
all_cibersort <- merge(all_cibersort, cell_analysis_MCCS[, !colnames(cell_analysis_MCCS) %in% c("P.value", "Correlation", "RMSE")], by = c("BMI", "BMI_Continuous", "Serology", "Serology_Continuous", "Day", "n_prev_vax", "Age"))

all_cibersort <- all_cibersort %>% mutate(Serology = factor(Serology),
                                                                    BMI = factor(BMI))

all_cibersort <- all_cibersort[,colnames(all_cibersort) %in% c(terms_to_plot, vars_3)] %>% arrange(Day, BMI, Serology)

mean_df_3 <- data.frame(matrix(nrow = 24))

for(i in 1:length(terms_to_plot)) {
  x <- all_cibersort %>%
    group_by_at(vars_3) %>%
        summarise_at(vars(terms_to_plot[[i]]),
                 list(Mean_Frequency = mean))
  mean_df_3[, terms_to_plot[[i]]] <- x$Mean_Frequency
  mean_df_3$Day <- x$Day
  mean_df_3$Serology <- x$Serology
  mean_df_3$BMI <- x$BMI
}

mean_df_2 <- data.frame(matrix(nrow = 24))

for(i in 1:length(terms_to_plot)) {
  x <- all_cibersort %>%
    group_by_at(vars_2) %>%
        summarise_at(vars(terms_to_plot[[i]]),
                 list(Mean_Frequency = mean))
  mean_df_2[, terms_to_plot[[i]]] <- x$Mean_Frequency
  mean_df_2$Day <- x$Day
  mean_df_2$Serology <- x$Serology
}

plotlist_3 <- list()
plotlist_2 <- list()

for(i in 1:length(terms_to_plot)){
  plot_df_3 <- mean_df_3[, c("Serology", "BMI", "Day", terms_to_plot[[i]])]
  plotlist_3[[i]] <- ggplot(data = plot_df_3, aes(x = !!plot_df_3[, "Day"], y = !!plot_df_3[, "Serology"], fill = !!plot_df_3[, terms_to_plot[[i]]])) +
  geom_tile() +
    labs(title = terms_to_plot[[i]]) +
  facet_grid(.~BMI)
  
    plot_df_2 <- mean_df_2[, c("Serology", "Day", terms_to_plot[[i]])]
  plotlist_2[[i]] <- ggplot(data = plot_df_2, aes(x = !!plot_df_2[, "Day"], y = !!plot_df_2[, "Serology"], fill = !!plot_df_2[, terms_to_plot[[i]]])) +
  geom_tile() +
    labs(title = terms_to_plot[[i]])
}

plotlist_2
plotlist_3
```


